% !TEX root =  ../main_manuscript.tex 

\begin{abstract}
\textbf{Background.} Low-risk prostate cancer patients enrolled in active surveillance (AS) programs commonly undergo biopsies for examination of cancer progression. Biopsies are conducted as per a fixed and frequent schedule (e.g. annual biopsies), common for all patients. Such schedules may schedule unnecessary biopsies. Since biopsies are burdensome, patients do not always comply with the schedule, which increases the risk of delayed detection of cancer progression.

\textbf{Objective.} Motivated by the world's largest AS program, Prostate Cancer Research International Active Surveillance (PRIAS), our aim is to better balance the number of biopsies (burden) and the delay in detection of cancer progression (benefit). We intend to achieve this by personalizing the decision of conducting biopsies.

\textbf{Methods.} Using joint models for time-to-event and longitudinal data, we jointly model the observed prostate-specific antigen levels, digital rectal examination scores, and the latest biopsy results of a patient at each follow-up visit. This results in a follow-up visit-specific and patient-specific, cumulative risk of cancer progression. When this risk at a visit is above a certain threshold, we schedule a biopsy at that visit. We compare this personalized approach with the currently practiced biopsy schedules via an extensive and realistic simulation study, based on a replica of the patients from the PRIAS program.

\textbf{Results.} In comparison to the currently practiced schedules, the personalized approach saves one to seven burdensome biopsies per patient, depending upon the time of cancer progression of the patient. Despite this reduction in the number of biopsies, the delay in the detection of cancer progression for the personalized approach remains comparable with that of the biopsy schedule of the PRIAS program.

\textbf{Conclusions.} We conclude that the personalized schedules better balance the number of biopsies per detected cancer progression.
\end{abstract}