% !TEX root =  ../main_manuscript.tex 
\begin{abstract}
\texttt{Background}: Prostate cancer active surveillance (AS) patients undergo repeat biopsies. Treatment commonly advised when biopsy Gleason~grade~$\geq$~2 (reclassification). Many patients never experience reclassification, yet undergo biopsies frequently. Reclassification risk based personalized biopsy schedules may reduce patient burden.\\

\texttt{Objective}: Develop web-application to assist patients/doctors make better biopsy decisions than fixed biopsies.\\

\texttt{Design, Setting, and Participants}: Model development: World's largest AS study PRIAS, 7813 patients, 1134 experienced reclassification; External validation: largest five GAP3 database cohorts; Data: prostate-specific antigen (PSA), repeat biopsy Gleason grade.\\

\texttt{Outcome Measurements, and Statistical Analysis}: Bayesian joint model fitted to PRIAS dataset. Model predicted patient-specific reclassification risk utilized for personalized biopsy decisions. Model validated in GAP3 cohorts using risk prediction error, calibration, area under ROC (AUC). Risk calculator, personalized schedules implemented in web-application, for PRIAS and validated GAP3 cohorts.\\

\texttt{Results and Limitations}: Reclassification rate at year five of follow-up: 35\% in PRIAS, at most 50\% in GAP3 cohorts. PSA velocity stronger predictor of reclassification (Hazard~Ratio:~2.47, 95\%CI:~1.93--2.99), than PSA value (Hazard~Ratio:~0.99, 95\%CI:~0.89--1.11). Validation: Moderate AUC (0.55--0.75) in PRIAS and GAP3 cohorts. Moderate prediction error (0.1--0.3) in GAP3 cohorts with Reclassification rate similar to PRIAS, large (0.3--0.45) otherwise. Model recalibrated for external GAP3 cohorts.\\

\texttt{Conclusions}: We successfully developed and validated web-application for predicting reclassification risks, and risk based personalized biopsy decisions, in prostate cancer AS. Available for PRIAS, and largest five GAP3 database cohorts. Enables shared decision making of biopsy schedules by comparing fixed and personalized schedules on total biopsies and expected time delay in detection of reclassification.\\

\texttt{Patient Summary}: Reclassification risk based personalized biopsy schedules are novel alternative to fixed biopsy schedules. They are implemented in our web-application. May offer better balance between total biopsies and time delay in detection of reclassification than fixed schedules.
\end{abstract}